---
title: "EDA-challenge"
author: "Minsoo Joo"
format: html
---

## EDA-challenge

```{r}
library(tidyverse)
```

```{r}
library(readr)
f <- "https://raw.githubusercontent.com/difiore/ada-datasets/main/data-wrangling.csv"
d <- read_csv(
  file = f,
  col_names = TRUE
)

names(d)
```

## Step 1

Create a new variable named BSD (body size dimorphism) which is the ratio of average male to female body mass.
Create a new variable named sex_ratio, which is the ratio of the number of adult females to adult males in a typical group.
Create a new variable named DI (for “defensibility index”), which is the ratio of day range length to the diameter of the home range.
```{r}
d <- d %>%
  mutate(
    BSD = Body_mass_male_mean / Body_mass_female_mean,
    sex_ratio = AdultFemale / AdultMales,
    DI = DayLength_km / (2 * sqrt(HomeRange_km2 / pi))
  )
```

Plot the relationship between day range length (y axis) and time spent moving (x axis), for these primate species overall and by family (i.e., a different plot for each family, e.g., by using faceting: + facet_wrap()). Do species that spend more time moving travel farther overall? How about within any particular primate family? Should you transform either of these variables?
```{r}
ggplot(d, aes(x = Move, y = DayLength_km)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~Family) +
  labs(x = "Time Spent Moving", y = "Day Range Length")
```

Plot the relationship between day range length (y axis) and group size (x axis), overall and by family. Do species that live in larger groups travel farther overall? How about within any particular primate family? Should you transform either of these variables?
```{r}
ggplot(d, aes(x = MeanGroupSize, y = DayLength_km)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ Family) +
  labs(x = "Group Size",
       y = "Day Range Length")
```

Plot the relationship between canine size dimorphism (y axis) and body size dimorphism (x axis) overall and by family. Do taxa with greater size dimorphism also show greater canine dimorphism?
```{r}
ggplot(d, aes(x = BSD, y = Canine_Dimorphism)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ Family) +
  labs(x = "Body Size Dimorphism",
       y = "Canine Size Dimorphism")
```

Create a new variable named diet_strategy that is “frugivore” if fruits make up >50% of the diet, “folivore” if leaves make up >50% of the diet, and “omnivore” if diet data are available, but neither of these is true (i.e., these values are not NA). Then, do boxplots of group size for species with different dietary strategies, omitting the category NA from your plot. Do frugivores live in larger groups than folivores?
```{r}
d <- d %>% 
  mutate(diet_strategy = case_when(
    Fruit >= 50 ~ "frugivore",
    Leaves >= 50 ~ "folivore",
    Fruit < 50 & Leaves < 50 ~ "omnivore",
    TRUE ~ NA
  ))

ggplot(filter(d, !is.na(diet_strategy)), aes(x = diet_strategy, y = MeanGroupSize)) +
  geom_boxplot() +
  labs(title = "group size for species with different dietary strategies",
       x = "diet_strategy",
       y = "MeanGroupSize")
```

In one line of code, using {dplyr} verbs and the forward pipe (|> or %>%) operator, do the following:
Add a variable, Binomial to the data frame d, which is a concatenation of the Genus and Species variables…
Trim the data frame to only include the variables Binomial, Family, Brain_size_species_mean, and Body_mass_male_mean…
Group these variables by Family…
Calculate the average value for Brain_Size_Species_Mean and Body_mass_male_mean per Family (remember, you may need to specify na.rm = TRUE)…
Arrange by increasing average brain size…
And print the output to the console
```{r}
d_summary <- d %>% 
  mutate(Binomial = paste(Genus, Species)) %>%
  select(Binomial, Family, Brain_Size_Species_Mean, Body_mass_male_mean) %>%
  group_by(Family) %>%
  summarise(avg_brain_size = mean(Brain_Size_Species_Mean, na.rm = TRUE),
            avg_body_mass = mean(Body_mass_male_mean, na.rm = TRUE)) %>%
  arrange(avg_brain_size)

print(d_summary)
```
